---
title: "Gamma"
author: "Aamina Thasneem Khaleel"
date: "2023-12-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Gamma Exposure 

Credit given to Toby Moran for gamma exposure calculations and chart development techniques for SPX

I added my own charts for AAPL and SOFI, and LCID. Since I have an interest in exploring trends within midcap stocks, I chose SOFI and LCID based on revenue and past earnings report.

Future Plans for the Project:
 - updating spot price for accuracy
 - pulling live data to report GEX charts in real-time
 - calculating gamma exposure based on expiration date

Looking for total put and call gamma at each strike  

total gamma at each strike

Libraries that I will be using

```{r}
library(data.table)
library(lubridate)
library(timeDate)
library(stringr)
library(formattable)
library(dplyr)
library(plotly)
library(tidyr)
library(purrr)
```
will be using fread which is faster and more convenient than read.table or read.csv

```{r}
#option chain data
optionChain <-fread("./spx_quotedata-2.csv")
lastSalePrices <- optionChain[, "Last Sale"]
lastSalePrices
spotLine <- fread("./spx_quotedata-2.csv", skip=3, nrows = 1)
spotLine
spotLineData <- colnames(spotLine[,3])
 # strsplit(as.character(colnames(spotLine[,3])), ":")
spotLineData
#obtained spot price for first row
spotPrice <- as.numeric(spotLine[1,3])
  #as.numeric(str_trim(spotLineData[[3]][1]))
spotPrice

#range of spot prices
fromS = 0.8*spotPrice
toS = 1.2*spotPrice
```
#Extracting current data of the spot price

```{r}
dateLine <- fread("./spx_quotedata-2.csv",skip =3, nrows = 1)
dateLine
todaysDateData <- strsplit(as.character(dateLine[,1]), " ")
# strsplit(as.character(colnames(dateLine[,1])), " ")
todaysDateData

todayYear <- todaysDateData[[1]][4]
todayMonth <- todaysDateData[[1]][2]
todayDay <- todaysDateData[[1]][3]
#using lubridate
todayDate <-as.Date(ymd(paste0(todayYear, '-', todayMonth, '-', todayDay)))
todayDate
```

Fixing the option chain column names
```{r}
colnames(optionChain) <-c('ExpirationDate','Calls','CallLastSale','CallNet','CallBid','CallAsk','CallVol',
'CallIV','CallDelta','CallGamma','CallOpenInt','StrikePrice','Puts','PutLastSale',
'PutNet','PutBid','PutAsk','PutVol','PutIV','PutDelta','PutGamma','PutOpenInt')
#changing the date format
optionChain$ExpirationDate <-as.Date(mdy(substring(optionChain$ExpirationDate,5)))

```

Will be calculating if the expiration is the third Friday of the month (monthly option expiration)

Note: SPX is for monthly expiration and SPXW is for weekly expiration

```{r}
#create another column
#need to use str split
expiration <- fread("./spx_quotedata-2.csv",skip =3, nrows = 1)
expiration
#expr <- strsplit(as.character(optionChain$Calls), "W")

expr <- sapply(strsplit(as.character(expiration[,2]), "W"), function(x) x[1])
sapply
expr
#using grepl and ifelse to determine monthly expiration
optionChain$monthlyExpiration <- ifelse(grepl("W", as.character(expiration[, 2])), FALSE, TRUE)

```
Calculating Gamma Spot Exposure

Option’s Gamma * Contract Size * Open Interest * Spot Price ^ 2 * 0.01

Call Gamma is positive

Put Gamma is negative

```{r}
optionChain$CallGEX <- optionChain$CallGamma * 100 * optionChain$CallOpenInt * spotPrice^2 *0.01
optionChain$PutGEX <- optionChain$PutGamma * 100 * optionChain$PutOpenInt * spotPrice^2 *0.01 * -1
optionChain$TotalGamma <- (optionChain$CallGEX + optionChain$PutGEX) / 10^9
#calculate the negative and pos
# strike (gamma * Open Interest * 100) and puts (gamma * Open Interest *-100)
#create a dataframe for plotting gex charts (both put, calls , and total gamma)
#range is +/ 20% of spot price
df <- optionChain %>% 
  filter(StrikePrice > fromS, StrikePrice < toS) %>%
  group_by(StrikePrice) %>% 
  summarise(CallGEX = sum(CallGEX), PutGEX = sum(PutGEX), TotalGamma = sum(TotalGamma))
#plotting variables
strike <- optionChain$StrikePrice
TotalGam <- round(sum(df$TotalGamma),3)
```

# Spot Gamma Exposure Chart

Total Gamma exposure over specified strike price range

```{r}
spotExpChart <- plot_ly(df,x= ~StrikePrice, y = ~TotalGamma, type = "bar", name = "Total Gamma")
spotExpChart <- spotExpChart %>% add_lines(x=spotPrice, line=list(color= "blue", dash = 'dot'), name= paste('Spot Price: ', round(spotPrice,0)))
spotExpChart <- spotExpChart %>%layout(
  title = paste("Total Gamma: $", TotalGam,  " Bn per 1% SPX Move ",  todayDate),
  xaxis=list(title='Strike Price'),
  yaxis=list(title='Spot Gamma Exposure ($ billions/1% move)')
)
spotExpChart
```

The graph above depicts the total gamma in billions per 1% of SPX move. There is a large amount of put gamma. 

Majority of calls are from covered calls (when an investor buys the underlying stock and sells the a call option). The market maker is long calls and short on the underlying stock. As price of the underlying stock increases, delta needs to be hedged and the market maker will sell.

Refer to this link to learn more about GEX (and its interpretation): https://help.quantdata.us/en/articles/7852449-what-is-gamma-exposure-gex

# Call and Put GEX Chart (for options)
```{r}
cpGEX <- plot_ly(data = df, x = ~StrikePrice, y = ~CallGEX, type = 'bar', name = 'Call Gamma')
cpGEX <- cpGEX %>% add_bars(
  x = ~StrikePrice,
  y = ~PutGEX,
  marker = list(color = 'red'),
  name = 'Put Gamma'
)
cpGEX <- cpGEX %>% add_lines(x=spotPrice,
                               line=list(color= 'blue', dash = 'dot'),
                               name= paste('Spot Price: ', round(spotPrice,0))
)
cpGEX <- cpGEX %>%layout(
  title = paste("Total Gamma: $", TotalGam,  " Bn per 1% SPX Move ",  todayDate),
  xaxis=list(title='Strike Price'),
  yaxis=list(title='Spot Gamma Exposure ($ billions/1% move)')
)
cpGEX
```

# Call Put Gex Charts for Stocks
## AAPL
```{r}
#obtain the last price
aaplOptChain <-fread("./aapl_quotedata.csv")
spotLine1 <- fread("./aapl_quotedata.csv", skip=3, nrows = 1)
spotPrice1 <- as.numeric((spotLine1[1,3]))
spotPrice1
#spotPrice range
fromS1 <- 0.8 * spotPrice1
toS1 <- 2.3 * spotPrice1
todaysDateData1 <-strsplit(as.character(spotLine1[,1]), " ")
todayYear1 <- todaysDateData1[[1]][4]
todayMonth1 <- todaysDateData1[[1]][2]
todayDay1 <- todaysDateData1[[1]][3]
todayDate1 <-as.Date(ymd(paste0(todayYear1, '-', todayMonth1, '-', todayDay1)))
todayDate1
colnames(aaplOptChain) <-c('ExpirationDate','Calls','CallLastSale','CallNet','CallBid','CallAsk','CallVol',
'CallIV','CallDelta','CallGamma','CallOpenInt','StrikePrice','Puts','PutLastSale',
'PutNet','PutBid','PutAsk','PutVol','PutIV','PutDelta','PutGamma','PutOpenInt')
aaplOptChain$CallGEX <- aaplOptChain$CallGamma * 100 * aaplOptChain$CallOpenInt * spotPrice1^2 *0.01
aaplOptChain$PutGEX <- aaplOptChain$PutGamma * 100 * aaplOptChain$PutOpenInt * spotPrice1^2 *0.01 * -1
aaplOptChain$TotalGamma <- (aaplOptChain$CallGEX + aaplOptChain$PutGEX) / 971
df1 <- aaplOptChain %>% 
  filter(StrikePrice > fromS1, StrikePrice < toS1) %>%
  group_by(StrikePrice) %>% 
  summarise(CallGEX = sum(CallGEX), PutGEX = sum(PutGEX), TotalGamma = sum(TotalGamma))
strike1 <- aaplOptChain$StrikePrice
TotalGam1 <- round(sum(df1$TotalGamma),3)
aaplSpotExpChart <- plot_ly(df1,x= ~df1$StrikePrice, y = ~df1$TotalGamma, type = "bar", name = "Total Gamma")
aaplSpotExpChart <- aaplSpotExpChart %>% add_lines(x=spotPrice1, line=list(color= "blue", dash = 'dot'), name= paste('Spot Price: ', round(spotPrice1,0)))
aaplSpotExpChart <- aaplSpotExpChart %>%layout(
  title = paste("Total Gamma: $", TotalGam1,  " AAPL",  todayDate1),
  xaxis=list(title='Strike Price'),
  yaxis=list(title='Spot Gamma Exposure ')
)
aaplSpotExpChart

aaplCPGEX <- plot_ly(data = df1, x = ~StrikePrice, y = ~CallGEX, type = 'bar', name = 'Call Gamma')
aaplCPGEX <- aaplCPGEX %>% add_bars(x = ~StrikePrice,
y = ~PutGEX, marker = list(color = 'red'), name = 'Put Gamma')
aaplCPGEX <- aaplCPGEX %>% add_lines(x=spotPrice1,
 line=list(color= 'blue', dash = 'dot'), name= paste('Spot Price: ', round(spotPrice1,0)))
aaplCPGEX <- aaplCPGEX %>%layout(title = paste("Total Gamma: $", TotalGam1,  "AAPL Move ",  todayDate1), xaxis=list(title='Strike Price'),yaxis=list(title='Spot Gamma Exposure'))
aaplCPGEX
```
Analysis: 

Interest in Small and Midcap stocks

# SOFI
```{r}
#obtain the last price
sofiOptChain <-fread("./sofi_quotedata.csv")
spotLine2 <- fread("./sofi_quotedata.csv", skip=3, nrows = 1)
spotPrice2 <- as.numeric((spotLine2[1,3]))
spotPrice2
#spotPrice range
fromS2 <- 0.7 * spotPrice2
toS2 <-2.5 * spotPrice2
todaysDateData2 <-strsplit(as.character(spotLine2[,1]), " ")
todayYear2 <- todaysDateData2[[1]][4]
todayMonth2 <- todaysDateData2[[1]][2]
todayDay2 <- todaysDateData2[[1]][3]
todayDate2 <-as.Date(ymd(paste0(todayYear2, '-', todayMonth2, '-', todayDay2)))
colnames(sofiOptChain) <-c('ExpirationDate','Calls','CallLastSale','CallNet','CallBid','CallAsk','CallVol',
'CallIV','CallDelta','CallGamma','CallOpenInt','StrikePrice','Puts','PutLastSale',
'PutNet','PutBid','PutAsk','PutVol','PutIV','PutDelta','PutGamma','PutOpenInt')
sofiOptChain$CallGEX <- sofiOptChain$CallGamma * 100 * sofiOptChain$CallOpenInt * spotPrice2^2 *0.01
sofiOptChain$PutGEX <- sofiOptChain$PutGamma * 100 * sofiOptChain$PutOpenInt * spotPrice2^2 *0.01 * -1
sofiOptChain$TotalGamma <- (sofiOptChain$CallGEX + sofiOptChain$PutGEX) / 327
df2 <- sofiOptChain %>% 
  filter(StrikePrice > fromS2, StrikePrice < toS2) %>%
  group_by(StrikePrice) %>% 
  summarise(CallGEX = sum(CallGEX), PutGEX = sum(PutGEX), TotalGamma = sum(TotalGamma))
strike2 <- sofiOptChain$StrikePrice
TotalGam2 <- round(sum(df2$TotalGamma),3)
sofiSpotExpChart <- plot_ly(df2,x= ~df2$StrikePrice, y = ~df2$TotalGamma, type = "bar", name = "Total Gamma")
sofiSpotExpChart <- sofiSpotExpChart %>% add_lines(x=spotPrice2, line=list(color= "blue", dash = 'dot'), name= paste('Spot Price: ', round(spotPrice2,0)))
sofiSpotExpChart <- sofiSpotExpChart %>%layout(
  title = paste("Total Gamma: $", TotalGam2,  " SOFI ",  todayDate2),
  xaxis=list(title='Strike Price'),
  yaxis=list(title='Spot Gamma Exposure ')
)
sofiSpotExpChart
sofiCPGEX <- plot_ly(data = df2, x = ~df2$StrikePrice, y = ~df2$CallGEX, type = 'bar', name = 'Call Gamma')
sofiCPGEX <- sofiCPGEX %>% add_bars(x = ~df2$StrikePrice,
y = ~df2$PutGEX, marker = list(color = 'red'), name = 'Put Gamma')
sofiCPGEX <- sofiCPGEX %>% add_lines(x=spotPrice2,
 line=list(color= 'blue', dash = 'dot'), name= paste('Spot Price: ', round(spotPrice2,0)))
sofiCPGEX <- sofiCPGEX %>%layout(title = paste("Total Gamma: $", TotalGam2,  "SOFI Move ",  todayDate2), xaxis=list(title='Strike Price'),yaxis=list(title='Spot Gamma Exposure'))
sofiCPGEX
```
Analysis:

# LCID
```{r}
#obtain the last price
lcidOptChain <-fread("./lcid_quotedata.csv")
spotLine3 <- fread("./lcid_quotedata.csv", skip=3, nrows = 1)
spotPrice3 <- as.numeric((spotLine3[1,3]))
spotPrice3
#spotPrice range
fromS3 <- 0.4 * spotPrice3
toS3 <- 2.8 * spotPrice3
todaysDateData3 <-strsplit(as.character(spotLine3[,1]), " ")
todayYear3 <- todaysDateData3[[1]][4]
todayMonth3 <- todaysDateData3[[1]][2]
todayDay3 <- todaysDateData3[[1]][3]
todayDate3 <-as.Date(ymd(paste0(todayYear3, '-', todayMonth3, '-', todayDay3)))
colnames(lcidOptChain) <-c('ExpirationDate','Calls','CallLastSale','CallNet','CallBid','CallAsk','CallVol',
'CallIV','CallDelta','CallGamma','CallOpenInt','StrikePrice','Puts','PutLastSale',
'PutNet','PutBid','PutAsk','PutVol','PutIV','PutDelta','PutGamma','PutOpenInt')
lcidOptChain$CallGEX <- lcidOptChain$CallGamma * 100 * lcidOptChain$CallOpenInt * spotPrice3^2 *0.01
lcidOptChain$PutGEX <- lcidOptChain$PutGamma * 100 * lcidOptChain$PutOpenInt * spotPrice3^2 *0.01 * -1
lcidOptChain$TotalGamma <- (lcidOptChain$CallGEX + lcidOptChain$PutGEX) / 345
df3 <- lcidOptChain %>% 
  filter(StrikePrice > fromS3, StrikePrice < toS3) %>%
  group_by(StrikePrice) %>% 
  summarise(CallGEX = sum(CallGEX), PutGEX = sum(PutGEX), TotalGamma = sum(TotalGamma))
strike3 <- lcidOptChain$StrikePrice
TotalGam3 <- round(sum(df3$TotalGamma),3)
lcidSpotExpChart <- plot_ly(df3,x= ~df3$StrikePrice, y = ~df3$TotalGamma, type = "bar", name = "Total Gamma")
lcidSpotExpChart <- lcidSpotExpChart %>% add_lines(x=spotPrice3, line=list(color= "blue", dash = 'dot'), name= paste('Spot Price: ', round(spotPrice3,0)))
lcidSpotExpChart <- lcidSpotExpChart %>%layout(
  title = paste("Total Gamma: $", TotalGam3,  " LCID ",  todayDate3),
  xaxis=list(title='Strike Price'),
  yaxis=list(title='Spot Gamma Exposure ')
)
lcidSpotExpChart
lcidCPGEX <- plot_ly(data = df3, x = ~df3$StrikePrice, y = ~df3$CallGEX, type = 'bar', name = 'Call Gamma')
lcidCPGEX <- lcidCPGEX %>% add_bars(x = ~df3$StrikePrice,
y = ~df3$PutGEX, marker = list(color = 'red'), name = 'Put Gamma')
lcidCPGEX <- lcidCPGEX %>% add_lines(x=spotPrice3,
 line=list(color= 'blue', dash = 'dot'), name= paste('Spot Price: ', round(spotPrice3,0)))
lcidCPGEX <- lcidCPGEX %>%layout(title = paste("Total Gamma: $", TotalGam2,  "LCID ",  todayDate3), xaxis=list(title='Strike Price'),yaxis=list(title='Spot Gamma Exposure'))
lcidCPGEX
```